\center \fontsize{18.8}{20} \selectfont Towards Globally Color-Dual All-Loop Integrands

### 2023-09-20 notes/2loopNLSM.tex
First read through

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\immediate\write18{mkdir -p images}  %% Create `pgf-img` directory
\usetikzlibrary{external}  
\immediate\write18{mkdir -p images}  %% Create `pgf-img` directory
\tikzexternalize[                     %% Activate externalization
 prefix=images/,                    %% Avoid cluttering the directory
 system call={                       %% Use lualatex in system call
    pdflatex \tikzexternalcheckshellescape -halt-on-error -shell-escape -interaction=batchmode -jobname="\image" "\texsource" || rm "\image.pdf"
 },
]
%Uncomment this  ^^^^^^^^ to compile in new images in real time. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

@@

%I think this belongs in another section...
%Why DC is cool:
%1) Reduction in computational complexity.
%Off-shell high spin is terrible.
%Gravity has infinitely many very complex Feynman rules as a result of gauge redundancy but YM only has two relatively simple Feynman rules.
%2) DC points to deep structure hidden in Lagrangian formulation.

Color-kinematics duality is a statement about the off-shell graphical
simplicity that manifest redundancy in the on-shell observables of
gauge and gravity theories. Ignoring factors of $i$ and the coupling
constant, a generic $m$-point $L$-loop amplitude in a color-dual
theory -- such as NLSM, YM, or bi-adjoint scalar
theory\ace{technically this is true as written for any theory.  In
  fact we slip that type of disclaimer in quite a lot throughout the
  rest of the section} -- can be written as
\begin{equation}
\mathcal{A}_m^L = \sum \limits_\Gamma \frac{1}{S_\Gamma} \int \frac{d^{LD}\ell}{(2\pi)^{LD}} \frac{c_\Gamma n_\Gamma}{d_\Gamma} .
\end{equation}
Here the spacetime dimension is $D$.  Even though the underlying
theory may have quartic (or higher multiplicity) vertices, it is
possible to multiply and divide by propagators so that the sum runs
over only purely cubic graphs $\Gamma$.  Overcounting is removed by
the internal symmetry factor $S_\Gamma$, which tracks the number of
graph symmetries with external legs held fixed.  The color factor
$c_\Gamma$ is the string of structure constants $f^{abc}$ associated
with the graph $\Gamma$,\footnote{This paper is focused on theories
  with particles that transform in the adjoint.  However, the double
  copy can be applied to a far more diverse set of theories with
  particles that transform in the fundamental or more diverse
  representations, see for example \cite{Johansson:2017srf,
    Johansson2014zca, Johansson:2015oia, Johansson:2019dnu,
    Carrasco:2023vjg, Carrasco:2020ywq}.}  the denominator $d_\Gamma$
is the product of propagators associated with the graph $\Gamma$, and
the numerator $n_\Gamma$ is all of the remaining kinematic factors
including dot products of momenta and polarization vectors.

The color factors $c_\Gamma$ are not independent and are related to
one another via a set of Jacobi identities that look schematically
like
\begin{equation}
c_i+c_j+c_k=0,
\end{equation}
where $i$, $j$, and $k$ are three graphs.  Example Jacobi relations
are presented in detail in \cref{sec:bootstrap}.  Color-kinematics
duality states that there exists a way to rearrange factors between
the various kinematic numerators $n_\Gamma$ such that the numerators
obey all of the same Jacobi identities as the color factors
\begin{equation}
n_i +n_j +n_k=0.
\end{equation}
If color-kinematics duality can be achieved at the off-shell level,
then the kinematic numerators should satisfy an algebra just like the
color factors.  In theories where the kinematic algebra is known, it
is often the algebra of volume preserving diffeomorphisms
\cite{Monteiro2011pc, Ben-Shahar:2021zww, Cheung:2020djz,
  Cheung:2021zvb, Cheung:2022mix}.  As a simple example, color-dual
numerators for 4pt pion scattering at tree level are
\begin{align}
&n_s = s(t-u)\\
&n_t = t(u-s)\\
&n_u = u(s-t)
\end{align}
where the coupling has been normalized away.  These numerators sum to
zero by explicit calculation.  Note that the $s$-channel numerator
contains an explicit factor of $s$ that cancels with the propagator
and the same goes for the $t$- and $u$-channels so that the net result
is a local function, just as one would expect for the pion tree
amplitude.

Once color-dual numerators $n_\Gamma$ have been found, they can be
used to replace the color factors ($c_\Gamma \to n_\Gamma$) in a
\emph{different} amplitude to produce the corresponding amplitude in
the double-copy theory \be\label{DCconstruction}
\tilde{\mathcal{A}}_m^L = \sum \limits_\Gamma \frac{1}{S_\Gamma} \int
\frac{d^{LD}\ell}{(2\pi)^{LD}} \frac{c_\Gamma
  \tilde{n}_\Gamma}{d_\Gamma} ~~ \xrightarrow{(c_\Gamma \to n_\Gamma)}
~~ \mathcal{M}_m^L = \sum \limits_\Gamma \frac{1}{S_\Gamma} \int
\frac{d^{LD}\ell}{(2\pi)^{LD}} \frac{n_\Gamma
  \tilde{n}_\Gamma}{d_\Gamma}.
\end{equation}
Importantly, the $\tilde{n}_\Gamma$ do not need to respect
color-kinematics duality.  The prototypical example of the double-copy
is that YM double copied with YM results in gravity where the gauge
invariance of each separate gluon produces the diffeomorphism
invariance of the graviton.  The double copy can be proven at tree
level using the KLT relations but there is significant evidence that
the double copy persists to all loop orders \cite{KLT, KiermaierTalk}
\citepls{Any loop papers we want?}.  In addition to pointing at deep
structure hidden in the Lagrangian for gravity, the double copy is
immensely useful at a utilitarian level since it is much more
efficient at producing $\mathcal{M}_m^L$ than traditional methods like
Feynman rules.

Many of the 1-loop results presented in \cref{sec:cubic} are informed
by the tree-level double-copy so we will review the tree case here.
The color structure of tree-level amplitudes is particularly simple.
A fully color dressed tree amplitude $\mathcal{A}_n$ in YM, NLSM, or
bi-adjoint scalar theory can be decomposed in terms of the trace basis
\begin{equation}
  \mathcal{A}_n = \sum \limits_{\sigma\in S_{n-1}} \text{Tr}(T^{a_{\sigma_1}} T^{a_{\sigma_2}}...T^{a_{\sigma_{n-1}}} T^{a_n}) A_n[\sigma_1, \sigma_2,...\sigma_{n-1}, n],
  \label{eq:trace-basis}
\end{equation}
where $T^a$ are generators of the gauge group and, because of the
cyclicity of the trace, one of the legs is held fixed so that the sum
runs over the $(n-1)!$ permutations of the remaining external legs.
The coefficients of the color factors are the color-ordered partial
amplitudes $A_n[...]$.  Any tree-level color factor can be converted
into a linear combination of Del Duca-Dixon-Maltoni (DDM) half-ladder
color factors
\begin{equation}
\text{DDM}[a_1, a_2,... a_n] \equiv f^{a_1 a_2 b_1} f^{b_1 a_3 b_2} f^{b_2 a_4 b_3}...f^{b_{n-3} a_{n-1} a_n}
\end{equation}
associated with the graph
\begin{equation}
\halfLadder
\end{equation}
by repeated application of the Jacobi identity \cite{DixonMaltoni}.
The fully color-dressed tree amplitude can then be re-expressed as
\begin{equation}
  \mathcal{A}_n = \sum \limits_{\sigma\in S_{n-2}} \text{DDM}[a_1, a_{\sigma_2}, a_{\sigma_3}, a_{\sigma_{n-1}}, a_n]  A_n[1, \sigma_2,...\sigma_{n-1}, n]
  \label{eq:ddm-basis}
\end{equation}
with legs $1$ and $n$ fixed so that the sum only contains $(n-2)!$
terms, indicating that there are additional relations -- the
Kleiss-Kuijf (KK) relations \cite{Kleiss:1988ne} -- amongst the
partial amplitudes appearing in the right-hand side of
\cref{eq:trace-basis}.  The KK relations are generic to any theory
with purely adjoint particles.  Color-kinematics duality further
implies that the KK basis is overcomplete and can be reduced to a
basis of $(n-3)!$ amplitudes via the fundamental BCJ identities
\cite{BCJ,Feng:2010my}
\begin{equation}
\sum \limits_{i=2}^{n-1} k_1 \cdot (k_2+...+k_i) A_n[2,...,i,1,i+1,...,n] =0.
\end{equation}
In terms of constructing explicit color-dual solutions at tree level,
it is enough to specify the numerator of the half ladder because any
diagram can be reduced to a half ladder through successive
applications of the Jacobi identity.  At 1-loop, every diagram can be
related to one topology as well, in this case the $n$-gon, but the
pattern does not persist to 2-loops.

\subsection{Color-dual Lagrangians}\label{offShellCK}
While the BCJ relations are an on-shell statement about the color-dual
nature of scattering amplitudes, one might aspire to hard code
color-kinematics duality by realizing an off-shell Lagrangian
description of the theory. The double-copy construction of
\cref{DCconstruction} would suggest a manifestly cubic Lagrangian of
the form
\begin{equation}
\mathcal{L} = P_V^{ab}P_W^{ij}\mathcal{O}^a_i \Box \mathcal{O}^b_j +  V^{abc} W^{ijk}\mathcal{O}^a_i \mathcal{O}^b_j \mathcal{O}^c_k
\end{equation}
where $V^{abc}$ and $W^{ijk}$ are cubic Feynman rules mixing field
operators $\mathcal{O}^{a}_i$ indexed by quantum numbers $a$ and
$i$. We have introduced $P_V^{ab}$ and $P_W^{ij}$ projection operators
to encode non-local kinetic structure that could in principle
participate in the construction.
The theory will be color dual if the cubic vertex is antisymmetric and satisfies the Jacobi identity, specifically,
\begin{align}
\textit{anti-symmetry}:& \quad{}_{a_1}\langle V^{a_2} \rangle_{a_3}+\text{cyc}(23)=0
\\
\textit{Jacobi identity}:& \quad {}_{a_1}\langle V^{a_2} V^{a_3}\rangle_{a_4}+\text{cyc}(234)=0
\end{align}
where 
\begin{equation}
{}_{a_1}\langle V^{a_2}V^{a_3}\cdots V^{a_{n-1}} \rangle_{a_n} \equiv V^{a_1a_2b_1}V^{b_1a_3b_2}\cdots V^{b_{n-3} a_{n-2}a_n}\,.
\end{equation}
Off-shell descriptions of
color-dual theories are rare, and only a few examples are available in
the literature \cite{Monteiro2011pc, Cheung:2016prv, Cheung:2021zvb,
  Cheung:2020djz, Ben-Shahar:2022ixa, Ben-Shahar:2021zww,
  Ben-Shahar:2021doh}.
\ace{Are these a reasonable collection?  Just grabbed the ones below}
\jm{These references get repeated in a second.  I'll fix this.}
Typically when one goes about constructing color-dual
theories directly from a set of cubic interactions, higher point
Jacobi and antisymmetry constraints require introducing additional
operators. See refs \cite{Ben-Shahar:2022ixa} for some recent
examples. Indeed, as we'll show in \cref{sec:cubic}, Yang-Mills
requires introducing a cubic 2-form interaction in order to manifest
the kinematic algebra just at four-point. Absent information about the
kinematic algebra, one can make progress in the construction of
loop-level color-dual amplitudes by employing a color-dual bootstrap.

@@

Aside from a few special theories that permit an off-shell Lagrangian
description \cite{Monteiro2011pc, Cheung:2016prv, Cheung:2021zvb,
  Cheung:2020djz, Ben-Shahar:2022ixa, Ben-Shahar:2021zww,
  Ben-Shahar:2021doh}, color-dual integrands are constructed from an
ansatz.  The four conditions we impose on an ansatz are summarized
here and elaborated on below.\footnote{The four ansatz conditions are
  similar to those of Ref.~\cite{Bern:2015ooa} except that there
  authors demand that the numerators have the same loop power counting
  as Feynman gauge Feynman rules.  This is an additional assumption
  that is most emphatically not made here.}
\begin{enumerate}
\item \textbf{Off-shell Locality}: Diagram numerators are polynomials
  in momenta and polarizations, \eg the diagrams only have poles given
  by the propagators of the diagrams.
\item \textbf{Color-kinematics duality}: The numerators obey the same
  Jacobi identities as the color factors.  This implicitly requires
  that the integrand is expressed in terms of purely cubic diagrams.
\item \textbf{Graph symmetries}: The diagram numerators are only a
  function of the diagram's topology and labeling.  Furthermore, the
  diagram numerators are invariant under the automorphisms of the
  diagrams \emph{including signs that compensate for color-factor sign
    changes}.
\item \textbf{On-shell Unitarity }: The cuts of the ansatz must
  reproduce the physical unitarity cuts of the theory when internal
  momenta are taken on-shell.
\end{enumerate}

In order to clarify general statements, the following subsections
include examples from a 4-point 2-loop integrand for some generic
color-dual theory like YM or NLSM.  This is the simplest case that
demonstrates the full machinery of the color-dual bootstrap.  All tree
and 1-loop processes are generated by single basis diagrams (the half
ladder and $n$-gon respectively) so they lack examples of the
``boundary'' Jacobi relations to be described shortly \ace{This is not
  strictly true.  Boundary Jacobi occur in 1L when demanding that BEL
  and/or tadpole are 0.  However, I appreciate that logic that 2L is
  the first time where an A+A+A=0 Jacobi occurs, rather than A+A=0.}.
Choosing the 4-point 2-loop integrand also has the advantage that this
process appears prominently in this paper.

@@

The kinematic numerators must be $D$-dimensional, Lorentz invariant,
local (polynomial) functions of the graph kinematics with the correct
power counting and external states.  Insisting on a local numerator is
partly a matter of convenience as the ansatz for a rational function
would quickly grow out of control without a guiding principle for what
terms to include.  Locality is also physically motivated since Feynman
rules typically produce polynomial numerators.  While these
assumptions are well motivated from a physical and complexity
standpoint, the literature has featured numerators that violate
locality, modify naive power counting through the inclusion of higher
spin modes, and break manifest Lorentz invariance \cite{Square,
  WeinzierlBCJLagrangian, Mogull:2015adi, FivePointN4BCJ,
  Johansson:2017bfl, Ben-Shahar:2022ixa, Cheung:2016prv,
  Cheung:2021zvb}.  These developments suggest that relaxing locality
is a viable alternative when the condition proves too stringent for an
ansatz.  We will comment more on locality in \cref{conclusions}.

@@

\paragraph{Color-kinematics duality}
\begin{figure}
\centering
{\MCutC, \MCutD, \sumoCut, \MCutJ,  
 \\
  \doubleBoxCut,  \crossBoxCut,\pentaTriangleCut,  \MCutF, 
\\
\MCutI,  \MCutE, \MCutK, 
\\
 \MCutL, \MCutG,  \MCutH,
}
\caption{The 14 non-pathological 4-point 2-loop cubic diagrams,
  including all dangling tree diagrams.} \label{fig:MaxCuts}
  \end{figure}

@@

Every numerator is required to respect all of the symmetries (or
automorphisms) of its associated graph.  Graph symmetries are an
avatar of Bose symmetry.  For a pure ansatz construction it is very
reasonable to impose graph symmetries, but color-kinematics duality at
tree-level is often achieved at the expense of manifest Bose symmetry
\cite{Cheung:2016prv, Cheung:2021zvb, Brandhuber:2021bsf,
  Edison:2020ehu}.  In an ansatz construction graph symmetries are
imposed for \emph{every} graph, not just the basis graphs.
%For example, the symmetries of the crossed box are spanned by
%\begin{equation}
%\CrossedBoxGraphSym{1}{2}{3}{4} \sim \CrossedBoxGraphSym{2}{1}{3}{4} \text{ and } \CrossedBoxGraphSym{1}{2}{3}{4} \sim \CrossedBoxGraphSym{2}{1}{4}{3}
% \end{equation}
\ace{is there a reason we commented out this diagram?}  For example,
the symmetries of the crossed box
\begin{equation}
\CrossedBoxGraphSym{1}{2}{3}{4}
\end{equation}
are spanned by $\{1234\} \to \{2134\}$ and $\{1234\} \to \{1243\}$.
In general, internal edge labels must be tracked as well.
Furthermore, care must be taken with the signs in a symmetry relation.
To compensate for the antisymmetric color factors, $f^{abc}$, graph
vertices are all totally antisymmetric and hence have signs built into
their orientations.  Once the signs are included in the symmetry
relations, the crossed box can be expressed in terms of the double box
and penta-triangle, thus imposing non-trivial constraints on the
ansatz. \ace{I'm not quite sure why the signs are mentioned in the
  previous sentence?}

@@

We now employ the color-dual bootstrap method described in \cref{bootstrap}. We will employ an ansatz constructed from  $D$-dimensionally Lorentz covariant kinematics. To fix remaining freedom in the NLSM bootstrap, we will match down to ZM theory in $D=2$. This is merely an aesthetic choice, so the final results of both the NLSM and Yang-Mills bootstraps are dimensionally agnostic.
\subsection{Two-loop NLSM}\label{2loopNLSM}

@@

The final essential constraint is unitarity.
Cuts (b) and (c) of \cref{eq:PotentialCuts} are the only physical cuts that need to be enforced.  \jm{Move some of the cut pictures here?}
This means that even at two loops, unitarity is still only probing the on-shell 4-point NLSM amplitude.
The other cuts besides (b) and (c) are either pathological or vanish because they involve a 3pt point on-shell amplitude (which vanishes on general grounds for scalar theories).
In order to manifest the $\mathbb{Z}_2$ symmetry of NLSM at the integrand level, we enforce that all non-pathological cuts containing a 3pt vertex (such as the maximal cut) vanish.
Imposing all relevant cuts leaves 765 free parameters representing the generalized gauge freedom in the answer.

@@

\subsection{Double-copy verification}\label{doubleCopyVerify}
\ace{I'm punting a lot of discussion to Ref \cite{Carrasco:2023qgz}.
  Do we think that is a problem?}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Comment the following out to compile new images
%
%\immediate\write18{mkdir -p images}  %% Create `pgf-img` directory
%\usetikzlibrary{external}  
%\immediate\write18{mkdir -p images}  %% Create `pgf-img` directory
%\tikzexternalize[                     %% Activate externalization
% prefix=images/,                    %% Avoid cluttering the directory
% system call={                       %% Use lualatex in system call
%    pdflatex \tikzexternalcheckshellescape -halt-on-error -shell-escape -interaction=batchmode -jobname="\image" "\texsource" || rm "\image.pdf"
% },
%]
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%Why DC is cool:
%1) Reduction in computational complexity.
%Off-shell high spin is terrible.
%Gravity has infinitely many very complex Feynman rules as a result of gauge redundancy but YM only has two relatively simple Feynman rules.
%2) DC points to deep structure hidden in Lagrangian formulation.

@@

Color-kinematics duality is a statement about the off-shell graphical
simplicity that manifest redundancy in the on-shell observables of
gauge and gravity theories. Ignoring factors of $i$ and the coupling
constant, a generic $m$-point $L$-loop amplitude in a color-dual
theory -- such as NLSM, YM, or bi-adjoint scalar
theory\ace{technically this is true as written for any theory.  In
  fact we slip that type of disclaimer in quite a lot throughout the
  rest of the section} \jm{I removed the second time we say `YM, NLSM, $\phi^3$' which was redundant.  I don't mind if we say it here because I want people to tell people that these are the kinds of theories we're thinking about.  We could omit it but then we'd have to say that the theory has at least one color factor and is probably an adjoint type theory and at that point I'd rather just give an example.} -- can be written as
\begin{equation}
\mathcal{A}_m^L = \sum \limits_\Gamma \frac{1}{S_\Gamma} \int \frac{d^{LD}\ell}{(2\pi)^{LD}} \frac{c_\Gamma n_\Gamma}{d_\Gamma} .
\end{equation}
Here the spacetime dimension is $D$.  Even though the underlying
theory may have quartic (or higher multiplicity) vertices, it is
possible to multiply and divide by propagators so that the sum runs
over only purely cubic graphs $\Gamma$.  Overcounting is removed by
the internal symmetry factor $S_\Gamma$, which tracks the number of
graph symmetries with external legs held fixed.  The color factor
$c_\Gamma$ is the string of structure constants $f^{abc}$ associated
with the graph $\Gamma$,\footnote{This paper is focused on theories
  with particles that transform in the adjoint.  However, the double
  copy can be applied to a far more diverse set of theories with
  particles that transform in the fundamental or more diverse
  representations, see for example \cite{Johansson:2017srf,
    Johansson2014zca, Johansson:2015oia, Johansson:2019dnu,
    Carrasco:2023vjg, Carrasco:2020ywq}.}  the denominator $d_\Gamma$
is the product of propagators associated with the graph $\Gamma$, and
the numerator $n_\Gamma$ is all of the remaining kinematic factors
including dot products of momenta and polarization vectors.

The color factors $c_\Gamma$ are not independent and are related to
one another via a set of Jacobi identities that look schematically
like
\begin{equation}
c_i+c_j+c_k=0,
\end{equation}
where $i$, $j$, and $k$ are three graphs.  Example Jacobi relations
are presented in detail in \cref{sec:bootstrap}.  Color-kinematics
duality states that there exists a way to rearrange factors between
the various kinematic numerators $n_\Gamma$ such that the numerators
obey the same Jacobi identities as the color factors
\begin{equation}
n_i +n_j +n_k=0.
\end{equation}
If color-kinematics duality can be achieved at the off-shell level,
then the kinematic numerators should satisfy an algebra just like the
color factors.  In theories where the kinematic algebra is known, it
is often the algebra of volume preserving diffeomorphisms
\cite{Monteiro2011pc, Ben-Shahar:2021zww, Cheung:2020djz,
  Cheung:2021zvb, Cheung:2022mix}.
When the kinematic algebra is not known explicitly, the kinematic numerators are typically constructed from an ansatz.
As an example, the result of such a calculation for 4pt pion scattering at tree level produces the numerators
\begin{align}
&n_s = s(t-u)\\
&n_t = t(u-s)\\
&n_u = u(s-t)
\end{align}
where the coupling has been normalized away.  These numerators sum to
zero by explicit calculation.  Note that the $s$-channel numerator
contains an explicit factor of $s$ that cancels with the propagator
and the same goes for the $t$- and $u$-channels so that the net result
is a local function, just as one would expect for the pion tree
amplitude.

Once color-dual numerators $n_\Gamma$ have been found, they can be
used to replace the color factors ($c_\Gamma \to n_\Gamma$) in a
\emph{different} amplitude to produce the corresponding amplitude in
the double-copy theory \be\label{DCconstruction}
\tilde{\mathcal{A}}_m^L = \sum \limits_\Gamma \frac{1}{S_\Gamma} \int
\frac{d^{LD}\ell}{(2\pi)^{LD}} \frac{c_\Gamma
  \tilde{n}_\Gamma}{d_\Gamma} ~~ \xrightarrow{(c_\Gamma \to n_\Gamma)}
~~ \mathcal{M}_m^L = \sum \limits_\Gamma \frac{1}{S_\Gamma} \int
\frac{d^{LD}\ell}{(2\pi)^{LD}} \frac{n_\Gamma
  \tilde{n}_\Gamma}{d_\Gamma}.
\end{equation}
Importantly, the $\tilde{n}_\Gamma$ do not need to respect
color-kinematics duality.  The prototypical example of the double-copy
is that YM double copied with YM results in gravity where the gauge
invariance of each separate gluon produces the diffeomorphism
invariance of the graviton.  The double copy can be proven at tree
level using the KLT relations but there is significant evidence that
the double copy persists to all loop orders \cite{FiveLoopN4, GeneralizedDoubleCopyFiveLoops, Bern:2018jmv, Carrasco:2021otn, KLT, KiermaierTalk}.
In addition to pointing at deep
structure hidden in the Lagrangian for gravity, the double copy is
immensely useful at a utilitarian level since it is much more
efficient at producing $\mathcal{M}_m^L$ than traditional methods like
Feynman rules.

Many of the 1-loop results presented in \cref{sec:cubic} are informed
by the tree-level double-copy so we will review the tree case here.
The color structure of tree-level amplitudes is particularly simple.
A fully color dressed tree amplitude $\mathcal{A}_n$ can be decomposed in terms of the trace basis
\begin{equation}
  \mathcal{A}_n = \sum \limits_{\sigma\in S_{n-1}} \text{Tr}(T^{a_{\sigma_1}} T^{a_{\sigma_2}}...T^{a_{\sigma_{n-1}}} T^{a_n}) A_n[\sigma_1, \sigma_2,...\sigma_{n-1}, n],
  \label{eq:trace-basis}
\end{equation}
where $T^a$ are generators of the gauge group and, because of the
cyclicity of the trace, one of the legs is held fixed so that the sum
runs over the $(n-1)!$ permutations of the remaining external legs.
The coefficients of the color factors are the color-ordered partial
amplitudes $A_n[...]$.  Any tree-level color factor can be converted
into a linear combination of Del Duca-Dixon-Maltoni (DDM) half-ladder
color factors
\begin{equation}
\text{DDM}[a_1, a_2,... a_n] \equiv f^{a_1 a_2 b_1} f^{b_1 a_3 b_2} f^{b_2 a_4 b_3}...f^{b_{n-3} a_{n-1} a_n}
\end{equation}
associated with the graph
\begin{equation}
\halfLadder
\end{equation}
by repeated application of the Jacobi identity \cite{DixonMaltoni}.
The fully color-dressed tree amplitude can then be re-expressed as
\begin{equation}
  \mathcal{A}_n = \sum \limits_{\sigma\in S_{n-2}} \text{DDM}[a_1, a_{\sigma_2}, a_{\sigma_3}, a_{\sigma_{n-1}}, a_n]  A_n[1, \sigma_2,...\sigma_{n-1}, n]
  \label{eq:ddm-basis}
\end{equation}
with legs $1$ and $n$ fixed so that the sum only contains $(n-2)!$
terms, indicating that there are additional relations -- the
Kleiss-Kuijf (KK) relations \cite{Kleiss:1988ne} -- amongst the
partial amplitudes appearing in the right-hand side of
\cref{eq:trace-basis}.  The KK relations are generic to any theory
with purely adjoint particles.  Color-kinematics duality further
implies that the KK basis is overcomplete and can be reduced to a
basis of $(n-3)!$ amplitudes via the fundamental BCJ identities
\cite{BCJ,Feng:2010my}
\begin{equation}
\sum \limits_{i=2}^{n-1} k_1 \cdot (k_2+...+k_i) A_n[2,...,i,1,i+1,...,n] =0.
\end{equation}
In terms of constructing explicit color-dual solutions at tree level,
it is enough to specify the numerator of the half ladder because any
diagram can be reduced to a half ladder through successive
applications of the Jacobi identity.  At 1-loop, every diagram can be
related to one topology as well, in this case the $n$-gon.
Two-loop calculations play an important role in our understanding of color-kinematics duality because this is the first order where multiple basis graphs are required, at least for a generic theory.

\subsection{Color-dual Lagrangians}\label{offShellCK}
While the BCJ relations are an on-shell statement about the color-dual
nature of scattering amplitudes, one might aspire to hard code
color-kinematics duality by realizing an off-shell Lagrangian
description of the theory. The double-copy construction of
\cref{DCconstruction} would suggest a manifestly cubic Lagrangian of
the form
\begin{equation}
\mathcal{L} = P_V^{ab}P_W^{ij}\mathcal{O}^a_i \Box \mathcal{O}^b_j +  V^{abc} W^{ijk}\mathcal{O}^a_i \mathcal{O}^b_j \mathcal{O}^c_k
\end{equation}
where $V^{abc}$ and $W^{ijk}$ are cubic Feynman rules mixing field
operators $\mathcal{O}^{a}_i$ indexed by quantum numbers $a$ and
$i$. We have introduced $P_V^{ab}$ and $P_W^{ij}$ projection operators
to encode non-local kinetic structure that could in principle
participate in the construction.
The theory will be color dual if the cubic vertex is antisymmetric and satisfies the Jacobi identity, specifically,
\begin{align}
\textit{anti-symmetry}:& \quad{}_{a_1}\langle V^{a_2} \rangle_{a_3}+\text{cyc}(23)=0
\\
\textit{Jacobi identity}:& \quad {}_{a_1}\langle V^{a_2} V^{a_3}\rangle_{a_4}+\text{cyc}(234)=0
\end{align}
where 
\begin{equation}
{}_{a_1}\langle V^{a_2}V^{a_3}\cdots V^{a_{n-1}} \rangle_{a_n} \equiv V^{a_1a_2b_1}V^{b_1a_3b_2}\cdots V^{b_{n-3} a_{n-2}a_n}\,.
\end{equation}
Off-shell descriptions of
color-dual theories are rare, and only a few examples are available in
the literature \cite{Monteiro2011pc, Cheung:2016prv, Cheung:2021zvb,
  Cheung:2020djz, Ben-Shahar:2022ixa, Ben-Shahar:2021zww,
  Ben-Shahar:2021doh}.
Typically when one goes about constructing color-dual
theories directly from a set of cubic interactions, higher point
Jacobi and antisymmetry constraints require introducing additional
operators. See ref \cite{Ben-Shahar:2022ixa} for a recent
example. Indeed, as we will show in \cref{sec:cubic}, Yang-Mills
requires introducing a cubic 2-form interaction in order to manifest
the kinematic algebra just at four-point. Absent information about the
kinematic algebra, one can make progress in the construction of
loop-level color-dual amplitudes by employing a color-dual bootstrap.

@@

When an off-shell color-dual Lagrangian is not known, the amplitude integrand must be constructed from an ansatz.  The four conditions we impose on an integrand ansatz are summarized
here and elaborated on below.\footnote{The four ansatz conditions are
  similar to those of Ref.~\cite{Bern:2015ooa} except that there the
  authors demand that the numerators have the same loop power counting
  as Feynman gauge Feynman rules.  This is an additional assumption
  that is most emphatically not made here.}
\begin{enumerate}
\item \textbf{Off-shell Locality}: Diagram numerators are polynomials
  in momenta and polarizations, e.g., the diagrams only have poles given
  by the propagators of the diagrams.
\item \textbf{Color-kinematics duality}: The numerators obey the same
  Jacobi identities as the color factors.  This implicitly requires
  that the integrand is expressed in terms of purely cubic diagrams.
\item \textbf{Graph symmetries}: The diagram numerators are only a
  function of the diagram's topology and labeling.  Furthermore, the
  diagram numerators are invariant under the automorphisms of the
  diagrams including signs that compensate for color-factor sign
    changes.
\item \textbf{On-shell Unitarity }: The cuts of the ansatz must
  reproduce the physical unitarity cuts of the theory when internal
  momenta are taken on-shell.
\end{enumerate}

In order to clarify general statements, the following subsections
include examples from a 4-point 2-loop integrand for some generic
color-dual theory like YM.  This is the simplest case that
demonstrates the full machinery of the color-dual bootstrap.  All tree
and 1-loop processes are generated by single basis diagrams (the half
ladder and $n$-gon respectively) so they lack examples of three-term
``boundary'' Jacobi relations to be described shortly.
Choosing the 4-point 2-loop integrand also has the advantage that this
process appears prominently in this paper.

@@

The kinematic numerators must be $D$-dimensional, Lorentz invariant,
local (polynomial) functions of the graph kinematics with the correct
power counting and external states.
Insisting on a local numerator is
partly a matter of convenience as the ansatz for a rational function
would quickly grow out of control without a guiding principle for what
terms to include.  Locality is also physically motivated since Feynman
rules typically produce polynomial numerators.  While these
assumptions are well motivated from a physical and complexity
standpoint, the literature has featured numerators that violate
locality, modify naive power counting through the inclusion of higher
spin modes, and break manifest Lorentz invariance \cite{Square,
  WeinzierlBCJLagrangian, Mogull:2015adi, FivePointN4BCJ,
  Johansson:2017bfl, Ben-Shahar:2022ixa, Cheung:2016prv,
  Cheung:2021zvb}.  These developments suggest that relaxing locality
is a viable alternative when the condition proves too stringent for an
ansatz.
In particular, the most natural way to guarantee locality is to require the off-shell numerators to be polynomial functions of kinematics.
However, simply requiring on-shell locality may suffice.
This alternative will be discussed in \cref{nonLocalScattering}.

@@

\paragraph{Color-kinematics duality}
\begin{figure}
\centering
{\MCutC, \MCutD, \sumoCut, \MCutJ,  
 \\
  \doubleBoxCut,  \crossBoxCut,\pentaTriangleCut,  \MCutF, 
\\
\MCutI,  \MCutE, \MCutK, 
\\
 \MCutL, \MCutG,  \MCutH
}
\caption{The 14 non-pathological 4-point 2-loop cubic diagrams,
  including all dangling tree diagrams.} \label{fig:MaxCuts}
  \end{figure}

@@

Every numerator is required to respect all of the symmetries (or
automorphisms) of its associated graph.  Graph symmetries are an
avatar of Bose symmetry.  For a pure ansatz construction it is very
reasonable to impose graph symmetries, but color-kinematics duality at
tree-level is often achieved at the expense of manifest Bose symmetry
\cite{Cheung:2016prv, Cheung:2021zvb, Brandhuber:2021bsf,
  Edison:2020ehu}.  In an ansatz construction graph symmetries are
imposed for \emph{every} graph, not just the basis graphs.
%For example, the symmetries of the crossed box are spanned by
%\begin{equation}
%\CrossedBoxGraphSym{1}{2}{3}{4} \sim \CrossedBoxGraphSym{2}{1}{3}{4} \text{ and } \CrossedBoxGraphSym{1}{2}{3}{4} \sim \CrossedBoxGraphSym{2}{1}{4}{3}
% \end{equation}
\ace{is there a reason we commented out this diagram?} \jm{You can try either one.  I liked the one I left in better.}
For example,
the symmetries of the crossed box
\begin{equation}
\CrossedBoxGraphSym{1}{2}{3}{4}
\end{equation}
are spanned by $\{1234\} \to \{2134\}$ and $\{1234\} \to \{1243\}$.
In general, internal edge labels must be tracked as well.
Furthermore, care must be taken with the signs in a symmetry relation.
To compensate for the antisymmetric color factors, $f^{abc}$, graph
vertices are all totally antisymmetric and hence have signs built into
their orientations.  Once the signs are included in the symmetry
relations, the crossed box can be expressed in terms of the double box
and penta-triangle, thus imposing non-trivial constraints on the
ansatz. \ace{I'm not quite sure why the signs are mentioned in the
  previous sentence?}

@@

We now employ the color-dual bootstrap method described in \cref{sec:bootstrap}. We will employ an ansatz constructed from  $D$-dimensionally Lorentz covariant kinematics. To fix remaining freedom in the NLSM bootstrap, we will match down to ZM theory in $D=2$. This is merely an aesthetic choice, so the final results of both the NLSM and Yang-Mills bootstraps are dimensionally agnostic.
\subsection{Two-loop NLSM}

@@

\begin{figure}
  \begin{center}
	\PhysicalCutOne{}{}{}{} \PhysicalCutTwo{}{}{}{}
  \end{center}
  \caption{The two physical cut topologies shared by NLSM, sGal, BI, and DBIVA}
  \label{fig:emu}
\end{figure}

The final essential constraint is unitarity.
The only two physical cuts that need to be enforced, shown in \cref{fig:emu}, only involve 4pt vertices meaning that even at two loops, unitarity is still only probing the on-shell 4-point NLSM amplitude.
Unitarity cuts besides those shown in \cref{fig:emu} are either pathological or vanish because they involve a 3pt point on-shell amplitude (which vanishes on general grounds for scalar theories).
In order to manifest the $\mathbb{Z}_2$ symmetry of NLSM at the integrand level, we require that all non-pathological cuts containing a 3pt vertex (such as the maximal cut) vanish.
Imposing all relevant cuts leaves 765 free parameters representing the generalized gauge freedom in the answer.

@@

\subsection{Double-copy verification}\label{doubleCopyVerify}
\ace{I'm punting a lot of discussion to Ref \cite{Carrasco:2023qgz}.
  Do we think that is a problem?} \jm{I think that's fine.}
